Semi-quantitative petrographic characterization of Bronze Age quartzite tools – a multivariate approach

### 2021-08-31 analysis/paper/paper.Rmd
Updated raman text and pca figures

The resulting PCA model (fig. \@ref(fig:nir-pca)) shows a clear trend along the first component (80.3 %) related to the above two features. The measurements with mainly quartz peaks gravitate towards the positive side of the 1st component, and those where the graphite peaks are dominant towards the opposite side. It is also clear that the darker material feature more pronounced graphite peaks, whereas the lighter and more translucent objects mainly feature quartz peaks. As in the case of the NIR data, however, there are exceptions that break this trend, as well as a “muddled” middle group with both well-pronounced graphite and quartz peaks. 
The 2nd component (6.82 %) mainly seems to differentiate spectra with a lot of noise, where the objects were difficult to sample. In contrast, along the 3rd component those samples that feature more pronounced peaks around 700 - 800 cm-1 groups up separately from the *quartz group*.

@@

```{r raman-pca, fig.width=18, fig.height=5, fig.fullwidth=TRUE, fig.cap="PCA score plot of 1st and 2nd components of the raman analysis. The materials group based on two main features, the quartz-peak at 460 cm-1 and the D/G band peaks around 1300 cm-1 and 1600 cm-1. Darker materials feature more pronounced D/G band peaks while the white and more translucent materials feature only the quartz peak."}
knitr::include_graphics('C:/Users/masj0062/Documents/DOKTORANDSPROJEKT/VERSION CONTROL/MethodPaper/analysis/figures/fig_4_3_nir_averaged_pca_p1_p2.jpg')
```

>>>>

The resulting PCA model (fig. \@ref(fig:nir-pca1)) shows a clear trend along the first component (81.5 %) related to the above two features. The measurements with mainly quartz peaks gravitate towards the positive side of the 1st component, and those where the graphite peaks are dominant towards the opposite side. It is also clear that the darker material feature more pronounced graphite peaks, whereas the lighter and more translucent objects mainly feature quartz peaks. As in the case of the NIR data, however, there are exceptions that break this trend, as well as a “muddled” middle group with both well-pronounced graphite and quartz peaks. 
The 2nd component (6.29 %) mainly seems to differentiate spectra with a lot of noise, where the objects were difficult to sample. In contrast, along the 3rd component those samples that feature more pronounced peaks around 700 - 800 cm-1 groups up separately from the *quartz group*.

@@

```{r raman-pca1, fig.width=18, fig.height=5, fig.fullwidth=TRUE, fig.cap="PCA score plot of 1st and 2nd components of the raman analysis. The materials group based on two main features, the quartz-peak at 460 cm-1 and the D/G band peaks around 1300 cm-1 and 1600 cm-1. Darker materials feature more pronounced D/G band peaks while the white and more translucent materials feature only the quartz peak."}
knitr::include_graphics('C:/Users/masj0062/Documents/DOKTORANDSPROJEKT/VERSION CONTROL/MethodPaper/analysis/figures/fig_4_6_raman_average_pca_p1_p2.jpg')
```

```{r raman-pca2, fig.width=18, fig.height=5, fig.fullwidth=TRUE, fig.cap="PCA score plot of 1st and 3rd components of the raman analysis. The materials group based on two main features, the quartz-peak at 460 cm-1 and the D/G band peaks around 1300 cm-1 and 1600 cm-1. Darker materials feature more pronounced D/G band peaks while the white and more translucent materials feature only the quartz peak."}
knitr::include_graphics('C:/Users/masj0062/Documents/DOKTORANDSPROJEKT/VERSION CONTROL/MethodPaper/analysis/figures/fig_4_7_raman_average_pca_p1_p3.jpg')
```
